\EL{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}

### 2020-02-27 jfm/draft_p2_jfm.tex
Iteration d'Emmanuel

From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
The use of analytical perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is unsuccessful.
Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
However, this requires very long simulations since these events are rare.
% general def on rare-event sampling %
%
{Rare-event sampling \EL{refers} to a large body of methods that aim at exploring preferentially regions of phase space corresponding to rare events, that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
%
In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} by using both very long time-series (direct sampling) and rare-event sampling techniques.
%

% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
%
In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}. In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
%
An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise). Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
%
However, a drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
%
%
%
In this paper, a more general approach is considered for complex, possibly deterministic, dynamical systems.
It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories, and is designed to sample rare events of some observable with a higher frequency.
%    
%
Even though such ideas date back to the early 1950s, they have received ever-growing interest over the the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
More importantly, these type of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
%
%
\EL{An original contribution of the present work is certainly to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom. 
%
Two algorithms a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.}
%

@@

The paper is organized in two parts. The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow. This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.
%
Motivation for this study is twofold.
Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations. This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible gain obtained from them. This assessment is developed in the second part of the paper.

% Annonce du plan
In more details,  the flow set-up is introduced and the dynamics related to typical
drag fluctuations is described in section~\ref{sec:test_flow}.
The statistical properties of the drag are then discussed.
In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme
	drag fluctuations} is investigated based on a direct sampling approach.
Both the instantaneous drag and time-averaged drag are considered.
It is found that sampled extreme events for the instantaneous drag share very similar dynamics. Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
%\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
Section~\ref{sec:rare_events_algorithms} reports the applicability of both the \ac{ams} and \ac{gktl} algorithms to the numerical simulation of extreme drag fluctuations, by using the same flow configuration.
In section~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly. This difficulty is put in perspective with the phenomenology developed in the previous sections.
Section~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag, using the \ac{gktl} algorithm.
This latter allows for an exceptional reduction of the computational cost needed to simulate trajectories corresponding to extreme values of the time-averaged drag.
As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
Perspectives and conclusion end this work.

@@

% lattice Boltzmann method 
%
The flow dynamics is integrated by the \ac{lbm} in our numerical simulations. 
While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at an underlying  mesoscopic level.
Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs. 
%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level. Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
In our context, this numerical method has been chosen principally for its computational manageability and efficiency.

% geometry
%
The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points. The square obstacle has size $R=16$ (in mesh unit) and is located at the centre of the channel. The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}). 
% boundary conditions
%
No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.  
%
Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively. The initial distributions are imposed at equilibrium (see Appendix \ref{app:lbm}). 
In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$. The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}. 
Near the end of the channel, the flow is progressively damped within a \textit{sponge layer} where the viscosity is artificially enhanced.
Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.

@@

% drag statistics
% pdf
Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
This long simulation will be referred to as the \textit{control run} in the following.
It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the lattice Boltzmann method. 
The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
%
Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
%
In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails. This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
\begin{figure}
	\centering
	\subfloat[\ac{pdf} of normalised drag fluctuations]
	{\label{fig:pdf_drag_a}
		\includegraphics[width=.45\linewidth]{PDF_drag/PDF_drag.png}}
	\subfloat[Autocorrelation of drag fluctuations]
	{\label{fig:pdf_drag_b}
		\includegraphics[width=.45\linewidth]{autocorrelation_drag/autocorrelation_drag.png}}  
	\caption{{(a)} \ac{pdf} of normalised drag fluctuations ($\tilde f_d = f_d - \bar{f_d}$) where $\mu = \bar{f_d}$ denotes the time-averaged value and $\sigma = \sqrt{\overline{{\tilde f_d}^2}}$ is the standard deviation. The drag is evaluated both in the presence (blue) and in the absence (red) of the obstacle. %Note that the amplitudes have not been normalized.
		\QQ{x-axis should be $x = (f_d-\mu)/\sigma)$ and y-axis $P(x)$)}
		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$. The correlation time \EL{$\tau_c\simeq 4 T_0$} is given by $C(\tau_c)=0$.
		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
	}
	\label{fig:pdf_drag}
\end{figure}

@@

How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
%
%
%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
one might expect that 
\begin{equation}
\label{eq:return_time}
\EL{r(a) \underset{a\to\infty}{\propto} \exp(la)}
\end{equation}
where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$. Let us point out some deviation from the exponential law at the largest levels, which are probably the consequence of under-sampling.

@@

% intro
%
We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
This set will be considered as representative of \emph{extreme events} in the upcoming study.
The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and possibly identify generic features.
%
According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma \equiv\sqrt{\overline{\tilde{f_d}^2}}$ being the standard deviation of the drag process.
Precisely, 104 independent fluctuations with $f_d(t) \geq 7.6\sigma$ have been identified. Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
%
In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.

@@

Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
\begin{equation}
\label{eq:avg_shear_def}
\overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
\end{equation}
where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
%
% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
%
Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does. 
This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values.

@@

% limitation of direct sampling
%
In the limit of very rare events and complex dynamics such as turbulent flows, the computational cost of  direct sampling becomes prohibitive. Indeed, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{l a}~(l>0)$ according to Eq.~\eqref{eq:return_time}.
As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges typically as $e^{l a}$.
%% objective
% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
The aim of rare-event sampling algorithms is to sample extreme fluctuations for a much lower computational cost.

\QQ{referee 3 asks for a general sentence about the physics behind these algorithms:}
\medskip
\QQ{Add a sentence to motivate the choice of these two algorithms for the present study. To what extent they are representative of rare-event sampling algorithms? Complementarity of the two algorithms? Make the link between fluid mechanics and statistical algorithms ?}
\medskip

\ZZ{Rare event algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$.
Then at each time step of the algorithm, some trajectories are discarded and others are replicated, in order to make trajectories with largest value of the extreme more common. The algorithm tracks the value of the ratio of the probability of the new ensemble with the probability of the previous one, in order to be able to estimate the probability of the extreme event in the model. Different algorithms use different selection rules. The success of the algorithm depends on the quality of the selection rule in order to select the precursors of the extreme event. Optimal selection rules are called committor functions, see for instance~\citep{lestang_computing_2018} and references therein. In the following we consider the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms. They both proved efficient for various dynamics and adopt opposite strategies:  the \acl{ams} algorithm uses as a selection rule the value of a predetermined score function, while the \acl{gktl} uses as a selection rule based on the increment of a score function.} \ZZ{A complete description of these two algorithms and their operating details is given in appendixes.}

%
%\ZZ{paragraph moved from introduction}
The \ac{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.  
In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}
or stochastic partial differential equations for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}. 
%
% The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}. 
 More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}. \ZZ{A review of the \ac{ams} algorithm, its history and applications is also available in \citep{cerou2019adaptive}}. 
%

>>>>

From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
The use of analytical perturbative methods in identifying resonant interactions (among degrees of freedom) responsible for extreme fluctuations is unsuccessful.
Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
However, this requires very long simulations since these extreme events are rare. \EL{The computational cost for sampling a fluctuation of very small probability typically grows as the inverse of this probability~\citep{wouters2016rare}.}
%
%However, this requires very long simulations since these events are rare.
% general def on rare-event sampling %
%
%
{Rare-event sampling \EL{refers} to a large body of methods that aim at \emph{preferentially} exploring  regions of phase space corresponding to  events that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
%
%	
% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
%
In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} by using both very long time-series (direct sampling) and rare-event sampling techniques.

%
In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}. In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
%
An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise). Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
%
However, a drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
%
%
%
Here, a more general approach is considered for complex, possibly deterministic, dynamical systems.
It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories of the system. 
%, and is designed to sample rare events of some observable with a higher frequency.
%    
%
Even though such ideas date back to the early 1950s, they have received ever-growing interest over the the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
More importantly, these types of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
%
%
\EL{An original contribution of our study is certainly to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom. 
%
Two algorithms a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.}
%

@@

The paper is organized in two parts. The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow. This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.
%
The motivation for this study is twofold.
Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations. This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible computational gain obtained from them. This assessment is developed in the second part of the paper.

% Annonce du plan
In more details,  the flow set-up is introduced and the dynamics related to typical
drag fluctuations is described in section~\ref{sec:test_flow}.
The statistical properties of the drag are then discussed.
In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme drag fluctuations} is investigated based on direct sampling.
Both the instantaneous drag and time-averaged drag are considered.
It is found in particular that sampled extreme events for the instantaneous drag result from very similar dynamics. %Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
%\sout{This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.}
Section~\ref{sec:rare_events_algorithms} examines the applicability of both the \ac{ams} and \ac{gktl} algorithms to the simulation of extreme drag fluctuations in the same flow configuration.
In subsection~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly. This difficulty is put in perspective with the phenomenology of extreme drag fluctuations developed in  section \ref{sec:direct_sampling}.
Subsection~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag by using the \ac{gktl} algorithm.
This latter allows for an exceptional reduction of the computational cost needed to simulate trajectories corresponding to extreme values of the time-averaged drag.
As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
Perspectives and conclusion end this work.

@@

% lattice Boltzmann method 
%
The flow dynamics is integrated numerically by the \ac{lbm} \EL{\citep{book_lbm}}. 
While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at an underlying  mesoscopic level.
Capturing the dynamics of collections of fluid particles moving and colliding on a lattice is here preferred to solving non-linear PDEs. 
%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level. Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
Further details about the \ac{lbm} are given in Appendix \ref{app:lbm} and references therein.
In our context, this numerical method has been chosen principally for its computational efficiency.

% geometry
%
The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points. The square obstacle has size $R=16$ (in mesh unit) and is located at the centre of the channel. The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}). 
% boundary conditions
%
No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\citep{lbm_book}.  
%
Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively. The initial distributions are imposed at equilibrium (see Appendix \ref{app:lbm}). 
In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$. The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}. 
Near the end of the channel, the flow is progressively damped within a \textit{sponge zone} where the viscosity is artificially enhanced.
Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.

@@

% drag statistics
% pdf
Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
This long simulation will be referred to as the \textit{control run} in the following.
It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the lattice Boltzmann method. 
The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
%
Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
%
In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails. This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
\begin{figure}
	\centering
	\subfloat[\ac{pdf} of normalised drag fluctuations]
	{\label{fig:pdf_drag_a}
		\includegraphics[width=.45\linewidth]{./PDF_drag/PDF_drag.png}}
	\subfloat[Autocorrelation of drag fluctuations]
	{\label{fig:pdf_drag_b}
		\includegraphics[width=.45\linewidth]{autocorrelation_drag/autocorrelation_drag.png}}  
	\caption{{(a)} \ac{pdf} of normalised drag fluctuations ($\tilde f_d = f_d - \bar{f_d}$) where $\mu = \bar{f_d}$ denotes the time-averaged value and $\sigma = \sqrt{\overline{{\tilde f_d}^2}}$ is the standard deviation. The drag is evaluated both in the presence (blue) and in the absence (red) of the obstacle. %Note that the amplitudes have not been normalized.
		\QQ{x-axis should be $x = (f_d-\mu)/\sigma)$ and y-axis $P(x)$)}
		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$. The correlation time \EL{$\tau_c\simeq 4 T_0$} is given by $C(\tau_c)=0$.
		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
	}
	\label{fig:pdf_drag}
\end{figure}

@@

How rare is a fluctuation $a$ is quantified by the return time $r(a)$.
We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
%
%
%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
one might expect that 
\begin{equation}
\label{eq:return_time}
\EL{r(a) \underset{a\to\infty}{\propto} \exp(la)}
\end{equation}
where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\citep{lestang_computing_2018}.
Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$. Let us also point out some deviation from the exponential law at the largest levels, which is probably the consequence of under-sampling.

@@

% intro
%
We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
This set will be considered as representative of \emph{extreme events} in the upcoming study.
The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and to possibly identify generic features.
%
According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma =\sqrt{\overline{\tilde{f_d}^2}}$ being the standard deviation of the drag process.
Precisely, 104 independent fluctuations with $f_d(t) \geq 7.6\sigma$ have been identified. Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
%
In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.

@@

Since the occurrence of large drag amplitudes arises from the production of vorticity along the top or bottom side of the square, it is proposed to characterize the dynamics of extreme events by their trajectory in the parameter space $(f_d(t), \bar{\gamma}(t))$ where $\bar{\gamma}(t)$ is the averaged shear along the top or bottom boundary of the square:
\begin{equation}
\label{eq:avg_shear_def}
\overline{\gamma} = \frac{1}{R} \int_{\mathcal{S}_\parallel} \frac{\partial u(\mathbf{x})}{\partial y}\mathrm{d}\mathbf{x},
\end{equation}
where $R$ denotes the size of the square, $u$ is the streamwise component of the velocity field and $\mathcal{S}_\parallel$ is the surface of either the top or the bottom boundary.
%
% Figure~\ref{fig:type_1_and_type_2_a} displays the label of the 104 sampled events, sorted according to the amplitude of the corresponding fluctuation.
% It illustrates that the dynamics corresponding to the large majority of events also corresponds to the events with the highest fluctuation amplitude.
%
Fig.~\ref{fig:shear_asof_drag} shows $\overline{\gamma}(t)$ as a function of the instantaneous drag $f_d$(t) for $t^{\star}-2\tau_c \leq t \leq  t^{\star}+2\tau_c$ for the 104 sampled extreme events.
Before and after the extremal fluctuation, \textit{i.e.} for $t^{\star}-2\tau_c \leq t \leq t^{\star}-\tau_c$ and $t^{\star}+\tau_c \leq t \leq t^{\star}+2\tau_c$, paths wander in a region related to typical values of both $\overline{\gamma}$ and $f_d$.
On the contrary, the drag abruptly varies for $t^{\star}-\tau_c \leq t \leq t^{\star}+\tau_c$ near the extremal amplitude.
%Paths in the $(f_d, \bar{\gamma})$ space display excursions to atypical values for both $\bar{\gamma}$ and $f_d$.
These excursions always go clockwise, that is, $\overline{\gamma}$ attains its maximum value before $f_d$ does. 
This is consistent with an increase of $\overline{\gamma}$ acting as a precursor for extreme drag fluctuations.
In this representation, we also observe that the path related to the increase of the drag is longer than the path related to the return to typical values\EL{, which may be related to the asymmetry of the mean profile displayed in Fig.~\ref{fig:timeseries_extremes}}.

@@

% limitation of direct sampling
%
In the limit of very rare events and complex dynamics such as turbulent flows, the computational cost of  direct sampling becomes prohibitive. Indeed, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{l a}~(l>0)$ according to Eq.~\eqref{eq:return_time}.
As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges typically as $e^{l a}$.
%% objective
% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
The aim of rare-event sampling algorithms is therefore to sample extreme fluctuations for a  computational cost much lower than their return time. 
%\QQ{referee 3 asks for a general sentence about the physics behind these algorithms:}
%\QQ{Add a sentence to motivate the choice of these two algorithms for the present study. To what extent they are representative of rare-event sampling algorithms? Complementarity of the two algorithms? Make the link between fluid mechanics and statistical algorithms ?}
%
\EL{Rare-event (sampling) algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, where $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the system.
At each step of the algorithm, certain trajectories are discarded and others are replicated in order to preferably sample trajectories with extreme fluctuations. The algorithm tracks the ratio of the probability of the new ensemble with the probability of the previous one, which makes it possible to estimate the statistical bias and, therefore, to infer the statistics of extreme events.

Different algorithms use different \emph{selection rules}. 
The success of the algorithm essentially depends on the quality of the selection rule to select the precursors of the extreme event \citep{rolland_statistical_2015}. 
However, with the exception of rare analytical cases, the optimal selection rules are not a priori known \citep{lestang_computing_2018}. 
In the following we consider the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms. They both proved to be efficient for various dynamics but adopt opposite strategies:  The \ac{ams} algorithm uses as a selection rule the value of a predetermined score function, while the \ac{gktl} uses a selection rule based on the increment of a score function.} \EL{A complete description of these two algorithms and their operating details are provided in the appendixes \ref{app:ams-short} and \ref{app:gktl_description}.}

%
%\ZZ{paragraph moved from introduction}
The \ac{ams} algorithm \citep{cerou_adaptive_2007} builds on previous ideas about splitting approaches \citep{KahnHarris1951,glasserman_multilevel_1999}.  
In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \citep{aristoff_adaptive_2015,teo_adaptive_2016}
or stochastic partial differential equations for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}. 
%
% The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \citep{rolland_computing_2016}. 
 More recently \EL{it} has been applied to rare events in stochastic models of wall-turbulence \citep{rolland_extremely_2018} and atmospheric dynamics \citep{bouchet2019rare}. \EL{A review of the \ac{ams} algorithm, its history and applications is also available in \citep{cerou2019adaptive}}. 
%
